Renal II

### 2017-05-28 renal2.rmd
typos for week ii

Pathophysiology: tl;dr glomerular filtration depends on pressure gradient from glomerulus to bowman space.  
Main determinant is renal blood flow (RBF). Reductions in RBF no matter the cause of AKI are a common pathological pathway for $\Downarrow$GFR. Prerenal = compromised renal perfusion. Intrinsic = usually glomerulus or tubule; ischaemic injury is the most common cause. Obstruction = increase in tubulur pressure decreases filtration driving force. This gradient equalises and renel efferent vasoconstriction. Depressed RBF leads to ischaemia and cell death. Inflammation comes in and cell injury occurs even after restoration. During this time tubular cell damage fucks up tight junctions => leakage of glom filtrate + dyaing cells enter tubules forming casts => decrease in GFR => oliguria.

Restoration of RBF  
Early RBF recovery has better prognosis. Once it is restored remaining nephrons increase filtration and hypertrophy, however if not enough nephrons left then these sclerose => more nephron loss. Hyperfiltration theory of renal failure: as nephrons die then puts more pressure on the rest which are then more prone to dying.

@@

Hypertension is a pre-disposing factor for => hypertrophy of wall => MI => blood leaking => cardiac tamponade

@@

Vegetations on heart valve - Rheumatic fever (acute immunologically mediated multisystem inflammatory disease that occurs a few weeks following an episode of group A Strep pharyngitis) leads to acute chronic rheumatic heart disease, which leads to chronic RHD (mitral stenosis commonest lesion) – fibrotic valvular disease – Left atrium dilates – congestion of lungs – Right ventricular hypertrophy (boot shaped heart) – dental extraction – infective endocarditis (bulky friable vegetations composed of thrombotic debris and organisma) – septicaemia and death.

@@

It is usually seen in debilitated patients eg. Cancer, where there is a hypercoaguable state which leads to ‘marantic vegetations’.

@@

Smoking => emphysema => centriacinar (not hereditary). $\alpha_{1}$-anti-trypsin would cause pan acinar

CXR: Flattening of the diaphragm => lung spaces have enalarged. Heart looks tubular (narrowing)

@@

Despite the “matched” V/Q defect, overtime hyperventilation develops and cardiac output (CO) drops which leads to areas of poor blood flow in relatively well oxygenated areas. Due to this poor CO, the rest of the body suffers from tissue hypoxia.

Pathogenesis of RHF: Matched V/Q defect --> Overtime hyperventilation develops --> Cardiac output drops --> Hypoxaemia --> Tissue hypoxia (this may cause polycythaemia)  and Increased CO2 retention (this may cause respiratory acidosis) --> Arterial vasoconstriction in lungs --> pulmonary hypertension --> Retrograde pressure build-up --> Elevated right ventricular pressure --> right ventricular hypertrophy --> right ventricular failure (Cor pulmonale)

@@

This x-ray shows a coin lesion in the right upper lobe.a

Hamartoma: excess tissue in a normal location

Choristoma: excess tissue in an abnormal location.

@@

-  Abcsess is an area of necrosis that is well sealed
-  Cavity: leaves behind a hole

@@

22F, fell from horse 1:45h ago, at rural hospital. Closed (within the skin; open means it breaks the skin and ready for osteomyolitis) spiral fracture of the 1/3 mid femur: neurovascular compromise suspected (scared of haematoma => compartment syndrome). Diastysis of pelvic symphysis. No haemoperitoneum. Circulating blood volume depletion estimated at 2L. Central venous line (so can pump lots in) and urinary catheter (measure urine output).

@@

$\0.5mL/kg/hr$ is the bare \textit{minimum} urine output.

If the vasa recta are not perfused adequately => ischaemia of the tubules => necrosis. It does not have a compensatory mechanism so if perfusion drops then it just cannot perfuse the tubular cells.

@@

-  Hyperkalaemia => ECG ASAP
-  Uraemia (increase in urea)

@@

Cystinosis => cysteine build up in proximal tubules => they start failing => lose electrolytes => Fanconi syndrome => decreased phosphate => decreased Ca absorption => rickles

@@

-  Kidney: pyelonephritis, tumour, trauma
-  Appendix: appendicitis
-  MSK: trauma, unlikely for haematuria
-  GIT: diverticulitis (diverticulum gets infected),  
-  Fistula
-  Could be referred pain (bladder: cystitis) that goes to kidneys

@@

PKD usually presents with pain and haematuria. Usually also have hypertension (due to fluid retention + activation of RAAS) and diastolic BP would increase. It causes stasis => kidney stone formation.

>>>>

Pathophysiology: tl;dr glomerular filtration depends on pressure gradient from glomerulus to bowmans space.  
Main determinant is renal blood flow (RBF). Reductions in RBF no matter the cause of AKI are a common pathological pathway for $\Downarrow$GFR. Prerenal = compromised renal perfusion. Intrinsic = usually glomerulus or tubule; ischaemic injury is the most common cause. Obstruction = increase in tubulur pressure decreases filtration driving force. This gradient equalises and renel efferent vasoconstriction. Depressed RBF leads to ischaemia and cell death. Inflammation comes in and cell injury occurs even after restoration. During this time tubular cell damage fucks up tight junctions $\rightarrow$ leakage of glom filtrate + dying cells enter tubules forming casts $\rightarrow$ decrease in GFR $\rightarrow$ oliguria.

Restoration of RBF  
Early RBF recovery has better prognosis. Once it is restored remaining nephrons increase filtration and hypertrophy, however if not enough nephrons left then these sclerose $\rightarrow$ more nephron loss. Hyperfiltration theory of renal failure: as nephrons die then puts more pressure on the rest which are then more prone to dying.

@@

Hypertension is a pre-disposing factor for $\rightarrow$ hypertrophy of wall $\rightarrow$ MI $\rightarrow$ blood leaking $\rightarrow$ cardiac tamponade

@@

Vegetations on heart valve $\rightarrow$ Rheumatic fever (acute immunologically mediated multisystem inflammatory disease that occurs a few weeks following an episode of group A Strep pharyngitis) leads to acute chronic rheumatic heart disease, which leads to chronic RHD (mitral stenosis commonest lesion) – fibrotic valvular disease $\rightarrow$  Left atrium dilates $\rightarrow$ congestion of lungs $\rightarrow$ Right ventricular hypertrophy (boot shaped heart) $\rightarrow$ dental extraction $\rightarrow$ infective endocarditis (bulky friable vegetations composed of thrombotic debris and organisma) $\rightarrow$ septicaemia and death.

@@

It is usually seen in debilitated patients eg. Cancer, where there is a hypercoaguable state which leads to 'marantic vegetations'.

@@

Smoking $\rightarrow$ emphysema $\rightarrow$ centriacinar (not hereditary). $\alpha_{1}$-anti-trypsin would cause pan acinar

CXR: Flattening of the diaphragm $\rightarrow$ lung spaces have enalarged. Heart looks tubular (narrowing)

@@

Despite the 'matched' V/Q defect, overtime hyperventilation develops and cardiac output (CO) drops which leads to areas of poor blood flow in relatively well oxygenated areas. Due to this poor CO, the rest of the body suffers from tissue hypoxia.

Pathogenesis of RHF: Matched V/Q defect $\rightarrow$ Overtime hyperventilation develops $\rightarrow$ Cardiac output drops $\rightarrow$ Hypoxaemia $\rightarrow$ Tissue hypoxia (this may cause polycythaemia)  and Increased CO2 retention (this may cause respiratory acidosis) $\rightarrow$ Arterial vasoconstriction in lungs $\rightarrow$ pulmonary hypertension $\rightarrow$ Retrograde pressure build-up $\rightarrow$ Elevated right ventricular pressure $\rightarrow$ right ventricular hypertrophy $\rightarrow$ right ventricular failure (Cor pulmonale)

@@

This x-ray shows a coin lesion in the right upper lobe.

-  Hamartoma: excess tissue in a normal location
-  Choristoma: excess tissue in an abnormal location.

@@

-  Abscess is an area of necrosis that is well sealed
-  Cavity: leaves behind a hole

@@

22F, fell from horse 1:45h ago, at rural hospital. Closed (within the skin; open means it breaks the skin and ready for osteomyolitis) spiral fracture of the 1/3 mid femur: neurovascular compromise suspected (scared of haematoma $\rightarrow$ compartment syndrome). Diastysis of pelvic symphysis. No haemoperitoneum. Circulating blood volume depletion estimated at 2L. Central venous line (so can pump lots in) and urinary catheter (measure urine output).

@@

$0.5mL/kg/hr$ is the bare *minimum* urine output.

If the vasa recta are not perfused adequately $\rightarrow$ ischaemia of the tubules $\rightarrow$ necrosis. It does not have a compensatory mechanism so if perfusion drops then it just cannot perfuse the tubular cells.

@@

-  Hyperkalaemia $\rightarrow$ ECG ASAP
-  Uraemia (increase in urea)

@@

Cystinosis $\rightarrow$ cysteine build up in proximal tubules $\rightarrow$ they start failing $\rightarrow$ lose electrolytes $\rightarrow$ Fanconi syndrome $\rightarrow$ decreased phosphate $\rightarrow$ decreased Ca absorption $\rightarrow$ rickles

@@

-  Kidney: pyelonephritis, tumour, trauma
-  Appendix: appendicitis
-  MSK: trauma, unlikely for haematuria
-  GIT: diverticulitis (diverticulum gets infected)
-  Fistula
-  Could be referred pain (bladder: cystitis) that goes to kidneys

@@

PKD usually presents with pain and haematuria. Usually also have hypertension (due to fluid retention + activation of RAAS) and diastolic BP would increase. It causes stasis $\rightarrow$ kidney stone formation.
